---
title: "Assignment2_Part1_VoiceInSchizophrenia"
author: "Riccardo Fusaroli"
date: "July 17, 2016"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
wd = setwd("C:/Users/annaa/OneDrive/3. semester/Exp. Method 3/Practical/Assignment4/Assignment4")
library(tseriesChaos);library(nonlinearTseries);library(SparseM);library(crqa);library(lme4);library(modelr);library(tibble);library(MuMIn);library(stringr);library(dplyr)


```

## Assignment 2 - Part 1 - Assessing voice in schizophrenia

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of diagnosis. Can we do better with our new fancy complex skills?

The corpus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). I have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

N.B. Question to be answered via email to Celine: can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interpret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.

N.B. There are looots of files to be dealt with. Probably too many for your computer. This is a challenge for you. Some (complementary) possible strategies: You can select a subset of files only (and you have to justify your choice). You can learn how to use the apply() or map() functions. You can coordinate with classmates.


###1.
In the course of this assignment you have to first select one datafile and figure out how to:

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)
- Extract "complex" descriptors: recurrence quantification analysis

```{r}
Pitch_file = read.delim("C:/Users/annaa/OneDrive/3. semester/Exp. Method 3/Practical/Assignment4/Assignment4/Pitch/Study1D0S101T2_f0.txt")
demo_data = read.delim("DemoData.txt")

#descriptive stats
mean(Pitch_file$f0) #139.92
sd(Pitch_file$f0) #31.38
median(Pitch_file$f0) #125.59
IQR(Pitch_file$f0) #42.14 intercortile range
mad(Pitch_file$f0) #23.77
sd(Pitch_file$f0)/mean(Pitch_file$f0)*100 #22.43 coefficient of variation

#RQA
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") #set parameter

#parameters
ans = optimizeParam(Pitch_file$f0,Pitch_file$f0, par,min.rec = 3.5, max.rec = 4.5) 
ans

Results = crqa(Pitch_file$f0, Pitch_file$f0, delay=35, embed = 3, radius = 5.72, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)


#Recurrence plot
RP=Results$RP
RP = matrix(as.numeric(RP), nrow = ncol(RP)) 
cols = c("white","blue4")
image(RP, xlab = "", ylab = "", col = cols)


```

###2. 
Second you will have to turn the code into a function and loop through all the files (or even better use apply/sapply/lapply)
- Remember to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)

```{r}

# make a list of the files that we are running through
files <- list.files(path = wd, pattern = "Study4", recursive = TRUE)
files

# make empty list for all the information of interest
Subject <- NULL
study <-  NULL
trial <-  NULL
diagnosis <-  NULL
PR = NULL
DET <- NULL
NRLINE <- NULL
maxL <- NULL
L <- NULL
ENTR <- NULL
LAM <- NULL
TT <- NULL
mean <- NULL
sd <- NULL
range <- NULL
median <- NULL
IQR = NULL
mad = NULL
coef_var = NULL

# for every loop n will be increased by 1 so that we can distinguish between each file
n = 1

# for loop to run through each file in the list (files) we have already created above
for (file in files) {
  
  # read the file into a data frame to be used in the loop
  df = read.delim(file)
  
  # create the values to be put in each of the empty lists that we created above
  Subject[n] <-  substr(file, 14, 18)
  study[n] <-  substr(file, 7, 12)
  trial[n] <- sub("_", "", substr(file, 20, 21))
  diagnosis[n] <-  substr(file, 14, 14)
  mean[n] <- mean(df$f0)
  sd[n] <- sd(df$f0)
  range[n] <- range(df$f0)
  median[n] <- median(df$f0)
  IQR[n] <- IQR(df$f0)
  mad[n] <- mad(df$f0,center = mean(df$f0))
  coef_var[n] <- sd(df$f0)/mean(df$f0)*100
  
  # set parameterlist and optimize them
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  
  parameters <- try(optimizeParam(df$f0, df$f0, par, min.rec = 3.5, max.rec = 4.5))
  
  
  # this if/else statement assures that we the files that does not fit the parameters are still considered in the final lists 
  if (length(parameters) > 1) {
  
  results1 <- crqa(df$f0, df$f0, delay = parameters$delay, embed = parameters$emddim, radius = parameters$radius, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)
  
  PR[n] = results1$RR
  DET[n] <- results1$DET
  NRLINE[n] <- results1$NRLINE
  maxL[n] <- results1$maxL
  L[n] <- results1$L
  ENTR[n] <- results1$ENTR
  LAM[n] <- results1$LAM
  TT[n] <- results1$TT
  
  
  }
  
  else {
  
  PR[n] = NA
  DET[n] <- NA
  NRLINE[n] <- NA
  maxL[n] <- NA
  L[n] <- NA
  ENTR[n] <- NA
  LAM[n] <- NA
  TT[n] <- NA
  
  }
  
#increase n by 1 to index into the next value in the lists
  n = n + 1
  
}

#safe it all in data frame
all_data_df_test <-  data.frame(Subject, study, trial, diagnosis, PR, DET, NRLINE, maxL, L, ENTR, LAM, TT, mean, sd, range, median)

#write a file for study 4 
write.csv(all_data_df_test, file = "dfstudy4")

#load the other studies 
study1 = read.csv("dfstudy1")
study2 = read.csv("dfstudy2")
study4 = read.csv("dfstudy4")

#merge the studies together 
recurrence_data = rbind(study1,study2,study4)
recurrence_data$Subject = substr(recurrence_data$Subject,3,5)

#incase of emergency
incaseofemergency = read.csv("final_rqa.csv")

```

###3. 

Make one model per acoustic feature and test whether you can observe significant difference due to Diagnosis. Tip: Which other fixed factors should you control for (that is, include in the model)? Which random ones?


- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 


```{r}

#models - whith acustic features features RQA

Model_range = lmerTest::lmer(PR ~ diagnosis + trial + (1+trial|study) + (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_DET = lmerTest::lmer(DET ~ diagnosis + trial + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_NRLINE = lmerTest::lmer(NRLINE ~ diagnosis + trial + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_maxL = lmerTest::lmer(maxL ~ diagnosis + trial + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_L = lmerTest::lmer(L ~ diagnosis + trial + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_ENTR = lmerTest::lmer(ENTR ~ diagnosis + trial + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_LAM = lmerTest::lmer(LAM ~ diagnosis + trial + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_TT = lmerTest::lmer(TT ~ diagnosis + trial + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)

summary(Model_DET)

#acustic features descriptive data
demo_data$Subject = substr(demo_data$Subject,3,5)
Gender_df = select(demo_data, 2,4)

Model_TT = lmerTest::lmer(mean ~ diagnosis + gender + trial + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)

#we can also use study as a random effect - cause researchers that run experiments do it diffenretly 
#If we didn't change the ID to be unique for controls and schitophrenic - we could have ut a random effect of = (1+Diagnosis+trial|Subject) - effect of diagnosis might vary across participant 


getmodel_information = function(mdl, n){
  temp_df = summary(mdl)$coefficients
  temp_df = as.data.frame(temp_df)
  temp_df = tibble::rownames_to_column(temp_df, "coef")
  temp_df$R2m = r.squaredGLMM(mdl)[1]
  temp_df$R2c = r.squaredGLMM(mdl)[2]
  temp_df$n = n
  return(temp_df)
}

Model_results_range = getmodel_information(Model_range, n = 1)
Model_results_DET = getmodel_information(Model_DET, n = 2)
#Model_results_NRLINE = getmodel_information(Model_NRLINE, n = 3)
Model_results_maxL = getmodel_information(Model_maxL, n = 4)
Model_results_L = getmodel_information(Model_L, n = 5)
Model_results_ENTR = getmodel_information(Model_ENTR, n = 6)
Model_results_LAM = getmodel_information(Model_LAM, n = 7)
Model_results_TT = getmodel_information(Model_TT, n = 8)



sejtdataframe = rbind(Model_results_range,Model_results_DET, Model_results_maxL, Model_results_L, Model_results_ENTR, Model_results_LAM, Model_results_TT)f

```


###3a. 
Is study a significant predictor in these models? What should you infer from this? Does study interact with diagnosis? What should you infer from this?



```{r}
Model_range2 = lmerTest::lmer(PR ~ diagnosis + trial + study + (1+trial|study) + (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_DET2 = lmerTest::lmer(DET ~ diagnosis + trial + study + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_NRLINE2 = lmerTest::lmer(NRLINE ~ diagnosis + trial + study + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_maxL2 = lmerTest::lmer(maxL ~ diagnosis + trial + study + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_L2 = lmerTest::lmer(L ~ diagnosis + trial + study + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_ENTR2 = lmerTest::lmer(ENTR ~ diagnosis + trial + study + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_LAM2 = lmerTest::lmer(LAM ~ diagnosis + trial + study + (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)
Model_TT2 = lmerTest::lmer(TT ~ diagnosis + trial + study (1+trial|study)+ (1 + diagnosis + trial|Subject), data =recurrence_data)

anova(Model_range,Model_range2)


```

4. Bonus Question: Compare effect size of diagnosis across the different measures. Which measure seems most sensitive?
- Tip: to compare across measures you need to put all of them on the same scale, that is, you need to "standardize" them (z-score)

```{r}


```

5. Bonus question. In the Clinical Info file you have additional information about the participants. Which additional parameters (e.g. age, gender) should we control for? Report the effects.

6. Write a paragraph reporting methods and results

[Next assignment: can we use these measures to build a tool that diagnoses people from voice only?]

## N.B. Remember to save the acoustic features of voice in a separate file, so to be able to load them next time


```{r}
#MERGE articulation (acustic feature) with demographic data
#read data
articulation = read.csv("Articulation.txt")
demo_data = read.delim("DemoData.txt")

#extract relevant info from string in subject
articulation$Subject=substr(articulation$soundname, 8, 12)

#set control to 0S and schizophrenia to 10 in demo data
demo_data$Diagnosis=revalue(demo_data$Diagnosis,c("Control"="0S","Schizophrenia"="1S"))

#Make coloumn with values in diagnosis and values in subject next to each other
demo_data$paste = paste(demo_data$Diagnosis,demo_data$Subject)

#delete the space 
demo_data$paste = sub(" ", "", demo_data$paste)                             

#call the "paste coloumn" "Subject"(the same as in articulation)
demo_data$Subject = demo_data$paste
                     
#merge demo_data and articulation data                           
data_all = merge(demo_data, articulation, by = "Subject")

data_all$Diagnosis=revalue(data_all$Diagnosis,c("0S"="Control","1S"="Schizophrenia"))

#make a trial coloumn
data_all$trial = substr(data_all$soundname, 14, 15) %>% as.numeric


```