---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

wd = setwd("C:/Users/annaa/OneDrive/3. semester/Exp. Method 3/Practical/Assignment4/Assignment4")
library(tseriesChaos);library(nonlinearTseries);library(SparseM);library(crqa);library(lme4);library(modelr);library(tibble);library(MuMIn);library(stringr);library(dplyr);library(caret);library(Metrics);library(pROC)

```

## Assignment 3 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.


### Question 1: Can you diagnose schizophrenia from pitch range only? If so, how well?

Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

```{r}
#load data and clean up a bit
df_features = read.csv("df_final")
df_features$X.1 = NULL
df_features$X = NULL
df_features$Subject = substr(df_features$Subject,3,5)

#model with random effects
log_model_range2 = glmer(diagnosis~ range + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))


df_features$Predictionslogodds = predict(log_model_range2)
df_features$Predictions[df_features$Predictionslogodds>0]="1"
df_features$Predictions[df_features$Predictionslogodds<=0]="0"

#calculate performance measures using full data set
confusionMatrix(data = df_features$Predictions, reference = df_features$diagnosis, positive = "0")

#rocCurve to calculate performance
rocCurve <- roc(response = df_features$diagnosis, predictor = df_features$Predictionslogodds)

auc(rocCurve)
#fair auc: 7.78
ci(rocCurve)
plot(rocCurve, legacy.axes = TRUE)

```

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?

```{r}

#Cross validation
df_features$Subject = as.factor(df_features$Subject)
df_features$Subject = as.numeric(df_features$Subject)


#Maybe use CVMS, allows you to have a list of models
#Cross validation

folds = createFolds(unique(df_features$Subject),k=5)

n=1

train_accuracy=NULL
train_sen = NULL
train_spec = NULL
train_npv = NULL
train_ppv = NULL
train_auc=NULL
train_ciLower=NULL
train_ciHigher=NULL

test_accuracy=NULL
test_sen = NULL
test_spec = NULL
test_npv = NULL
test_ppv = NULL
test_auc=NULL
test_ciLower=NULL
test_ciHigher=NULL


for (i in folds)
  {
  testD=df_features[df_features$Subject %in% i,]
  trainD=df_features[!(df_features$Subject %in% i),]
  Mtrain = lme4::glmer(diagnosis~range + trial + (1+trial|Subject), family = "binomial", data = trainD, control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
  
trainD$Predictionslogodds = predict(Mtrain,trainD,allow.new.levels=TRUE)
trainD$Predictions[trainD$Predictionslogodds>0] = "1"
trainD$Predictions[trainD$Predictionslogodds<0] = "0"
train_matrix = confusionMatrix(data = trainD$Predictions,reference = trainD$diagnosis, positive = "0")

train_accuracy[n]=train_matrix$overall[1]
train_sen[n]=train_matrix$byClass[1]
train_spec[n]=train_matrix$byClass[2]
train_ppv[n]=train_matrix$byClass[3]
train_npv[n]=train_matrix$byClass[4]

testD$Predictionslogodds = predict(Mtrain,testD,allow.new.levels=TRUE)
testD$Predictions[testD$Predictionslogodds>0] = "1"
testD$Predictions[testD$Predictionslogodds<0] = "0"
test_matrix = confusionMatrix(data = testD$Predictions,reference = testD$diagnosis, positive = "0")

test_accuracy[n]=train_matrix$overall[1]
test_sen[n]=test_matrix$byClass[1]
test_spec[n]=test_matrix$byClass[2]
test_ppv[n]=test_matrix$byClass[3]
test_npv[n]=test_matrix$byClass[4]

train_RocCurve=roc(response=trainD$diagnosis,predictor = trainD$Predictionslogodds)
train_auc[n]=train_RocCurve$auc
train_ciLower[n]=ci(train_RocCurve)[1]
train_ciHigher[n]=ci(train_RocCurve)[2]

test_RocCurve=roc(response=testD$diagnosis,predictor = testD$Predictionslogodds)
test_auc[n]=test_RocCurve$auc
test_ciLower[n]=ci(test_RocCurve)[1]
test_ciHigher[n]=ci(test_RocCurve)[2]

  n=n+1
  
}
traindf = data.frame(train_accuracy,train_sen,train_spec,train_npv,train_ppv,train_auc,train_ciLower,train_ciHigher)

testdf = data.frame(test_accuracy,test_sen,test_spec,test_npv,test_ppv,test_auc,test_ciLower,test_ciHigher)




```


### Question 2 - Which single acoustic predictor is the best predictor of diagnosis?


```{r}

#i = predictor, x = data, 

cvalid = function(i, p){

test_accuracy=NULL
test_sen = NULL
test_spec = NULL
test_npv = NULL
test_ppv = NULL
test_auc=NULL
test_ciLower=NULL
test_ciHigher=NULL

n=1

for (fold in folds)
  {
  testD=df_features[df_features$Subject %in% fold,]
  trainD=df_features[!(df_features$Subject %in% fold),]
  Mtrain = i

testD$Predictionslogodds = predict(Mtrain,testD, allow.new.levels=TRUE)
testD$Predictions[testD$Predictionslogodds>0] = "1"
testD$Predictions[testD$Predictionslogodds<0] = "0"
test_matrix = confusionMatrix(data = testD$Predictions,reference = testD$diagnosis, positive = "0")

test_accuracy[n]=test_matrix$overall[1]
test_sen[n]=test_matrix$byClass[1]
test_spec[n]=test_matrix$byClass[2]
test_ppv[n]=test_matrix$byClass[3]
test_npv[n]=test_matrix$byClass[4]

test_RocCurve=roc(response=testD$diagnosis,predictor = testD$Predictionslogodds)
test_auc[n]=test_RocCurve$auc
test_ciLower[n]=ci(test_RocCurve)[1]
test_ciHigher[n]=ci(test_RocCurve)[2]

  n=n+1
  
}
testdf = data.frame(mean(test_accuracy),mean(test_sen),mean(test_spec),mean(test_npv),mean(test_ppv),mean(test_auc),mean(test_ciLower),mean(test_ciHigher))

return(testdf)
}

#make models with different predictors
log_model_sd = glmer(diagnosis~ sd + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_PR = glmer(diagnosis~ PR + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_DET = glmer(diagnosis~ DET + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_NRLINE = glmer(diagnosis~ NRLINE + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_maxL = glmer(diagnosis~ maxL + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_L = glmer(diagnosis~ L + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_ENTR = glmer(diagnosis~ LAM + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_TT = glmer(diagnosis~ TT + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_mean = glmer(diagnosis~ mean + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_range = glmer(diagnosis~ range + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
log_model_median = glmer(diagnosis~ median + trial + (1+trial|Subject), data = df_features, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

#data frames for each predictor
a = cvalid(log_model_sd)
a$predictor = "sd"

b = cvalid(log_model_PR)
b$predictor = "PR"

c = cvalid(log_model_DET)
c$predictor = "DET"

d = cvalid(log_model_NRLINE)
d$predictor = "NRLINE"

e = cvalid(log_model_maxL)
e$predictor = "maxL"

f = cvalid(log_model_L)
f$predictor = "L"

g = cvalid(log_model_ENTR)
g$predictor = "ENTR"

h = cvalid(log_model_TT)
h$predictor = "TT"

i = cvalid(log_model_mean)
i$predictor = "mean"

j = cvalid(log_model_range)
j$predictor = "range"

#data frame with performance measure auc for all the predictors
all_predictorsdf = rbind(a,b,c,d,e,f,g,h,i,j)

```


### Question 3 - Which combination of acoustic predictors is best for diagnosing schizophrenia?

Now it's time to go wild! Use all (voice-related) variables and interactions you can think of. Compare models and select the best performing model you can find.

Remember:
- Out-of-sample error crucial to build the best model!
- After choosing the model, send Celine and Riccardo the code of your model

```{r}



```

### Question 4: Properly report the results

METHODS SECTION: how did you analyse the data? That is, how did you extract the data, designed the models and compared their performance?

RESULTS SECTION: can you diagnose schizophrenia based on voice? which features are used? Comment on the difference between the different performance measures.

### Bonus question 5

You have some additional bonus data involving speech rate, pauses, etc. Include them in your analysis. Do they improve classification?

### Bonus question 6

Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them.
